
---
title: "Selecting Subjects for PNC Template"
author: "BBL, et al"
date: "October 10, 2015"
output: pdf_document
---

*Selecting Subjects for PNC Template*
=========================


Goal here is to select a relatively matched sample (n=120) for the PNC template.

Data pulled from the July 30 PNC data release on Galton:

/home/analysis/redcap_data/201507//n1601_go1_datarel_073015.csv

_Stratgegy is as follows:_

1.  Remove subjects who failied visual QA (could be debated) & health problems
2.  Divide into 5 age bins (quantiles)
3.  In each age bin, select 12 subjects who are healthy (no psychopathology/meds) and 12 who are not healthy (any psychopathology, meds ok)
4.  For each of those 12 subjects , evenly matched on sex*race


```{r, echo=FALSE,eval=TRUE,warning=FALSE}
library(knitr)

#load data
go1Name<-"/Users/sattertt/Documents/Magic Briefcase/ACTIVE_PROTOCOLS/EONS/1601/dataReleases/n1601_go1_datarel_073015.csv"
data<-read.csv(go1Name)

#fix certain variables
data$male<-NA
data$male[which(data$sex==2)]<-"female"
data$male[which(data$sex==1)]<-"male"
data$male<-as.factor(data$male)

data$age<-data$ageAtGo1Scan/12

data$white<-NA
data$white[which(data$race==1)]<-"caucasian"
data$white[which(data$race!=1)]<-"notCaucaisian"
data$white<-as.factor(data$white)

data$healthy<-"notHealthy"
data$healthy[which(data$goassessSmryPsychOverallRtg<4 & data$goassessPstd!="4PS" & data$ltnExclude==0)]<-"healthy"
data$health<-as.factor(data$healthy)

#create age bin quantiles
ageBins<-quantile(data$age,probs = seq(0, 1, 0.20))
data$ageBin<-NA
for (j in 1:length(ageBins)){
  data$ageBin[which(data$age>=ageBins[j] & data$age<ageBins[j+1] )]<-j
}
data$ageBin[which(data$age==ageBins[6])]<-5

#remove subjects who failed visual QA or have health problems
dataQa<-data[which(data$mprageSbiaExclude==0 & data$healthExclude==0),c("bblid","scanid","age","male","white","ageBin","healthy")]

#goal is to have 60 healthy, 60 non healthy
#10 each per quantile, 6 male, 6 female, 3 whtie, 3 non-white
templateSubj<-matrix(nrow=120,ncol=1,NA)
rowStart=1
for (i in 1:max(data$ageBin)){
    idsTmp<-NA
    rowStart<-(i-1)*24+1
    rowEnd<-rowStart+23
  #  print(rowStart)
   # print(rowEnd)
    
    #no healthy
    idsTmp<-which(dataQa$ageBin==i & dataQa$healthy=="notHealthy" & dataQa$male=="male" & dataQa$white=="caucasian")
    ids1<-dataQa$bblid[idsTmp[1:3]]
    
    idsTmp<-which(dataQa$ageBin==i & dataQa$healthy=="notHealthy" & dataQa$male=="male"  & dataQa$white!="caucasian")
    ids2<-dataQa$bblid[idsTmp[1:3]]
    
    idsTmp<-which(dataQa$ageBin==i & dataQa$healthy=="notHealthy" & dataQa$male=="female" & dataQa$white=="caucasian")
    ids3<-dataQa$bblid[idsTmp[1:3]]
    
    idsTmp<-which(dataQa$ageBin==i & dataQa$healthy=="notHealthy" & dataQa$male=="female" & dataQa$white!="caucasian")
    ids4<-dataQa$bblid[idsTmp[1:3]]
    
    #healthy
    
    idsTmp<-which(dataQa$ageBin==i & dataQa$healthy=="healthy" & dataQa$male=="male"  & dataQa$white=="caucasian")
    ids5<-dataQa$bblid[idsTmp[1:3]]
    
    idsTmp<-which(dataQa$ageBin==i & dataQa$healthy=="healthy"  & dataQa$male=="male"  & dataQa$white!="caucasian")
    ids6<-dataQa$bblid[idsTmp[1:3]]
    
    idsTmp<-which(dataQa$ageBin==i & dataQa$healthy=="healthy"  & dataQa$male=="female" & dataQa$white=="caucasian")
    ids7<-dataQa$bblid[idsTmp[1:3]]
    
    idsTmp<-which(dataQa$ageBin==i & dataQa$healthy=="healthy"  & dataQa$male=="female" & dataQa$white!="caucasian")
    ids8<-dataQa$bblid[idsTmp[1:3]]
    
    idComb<-c(ids1,ids2,ids3,ids4,ids5,ids6,ids7,ids8)
    
    
    templateSubj[rowStart:rowEnd]<-idComb
}

templateSubjData<-dataQa[which(dataQa$bblid %in% templateSubj),]
```


=========================

### Now sanity checking to make sure this worked as planned . . . .here divided by age bins

```{r, echo=FALSE}
#tables for sanity checking
kable(table(templateSubjData$ageBin,templateSubjData$male))
kable(table(templateSubjData$ageBin,templateSubjData$white))
kable(table(templateSubjData$ageBin,templateSubjData$healthy))
kable(table(templateSubjData$male,templateSubjData$healthy))
kable(table(templateSubjData$white,templateSubjData$healthy))
kable(table(templateSubjData$white,templateSubjData$male))
```

=========================

### Check to see if age is similar across the different divisions. . . . 
````{r, echo=FALSE}
#tables for sanity checking
ageMeanSex<-tapply(templateSubjData$age,templateSubjData$male,mean)
kable(as.data.frame(ageMeanSex))

ageMeanHealth<-tapply(templateSubjData$age,templateSubjData$healthy,mean)
kable(as.data.frame(ageMeanHealth))

ageMeanRace<-tapply(templateSubjData$age,templateSubjData$white,mean)
kable(as.data.frame(ageMeanRace))

```

=========================

### Finally check that age is reasonably balanced over the range 

````{r, echo=FALSE}
hist(templateSubjData$age)
```



